---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r setoptions, echo=TRUE}
Sys.setlocale(category = 'LC_ALL', locale='en_US.UTF-8')

# unzip activity.zip, if not already unzipped
if (!file.exists('activity.csv'))
    unzip(zipfile = 'activity.zip', exdir = '.')

# read data
data <- read.csv(file = 'activity.csv', header = TRUE, sep = ',')
data$date <- as.Date(as.character(data$date), '%Y-%m-%d')
tidy <- data[!is.na(data$steps), ]
```


## What is mean total number of steps taken per day?
```{r echo=TRUE}
d <- aggregate(steps~date, tidy, sum)
plot(d$date, d$steps, type="h", ylab = 'Total steps', xlab = 'Date', 
     main = 'Total steps in 2012')
```

- The *mean* total number of steps taken per day is **`r mean(d$steps, na.rm = T)`**
- The *median* total number of steps taken per day: **`r median(d$steps, na.rm = T)`**


## What is the average daily activity pattern?
```{r echo=TRUE}
dinterval <- aggregate(steps~interval, tidy, mean)
plot(dinterval, type='l', ylab='Average number of steps taken', xlab='interval of 5 minutes')

# find max index
index <- which.max(d$steps)
```
Maximun 5-minute interval is entry number **`r index`** with values **interval `r d[index, 1]`** and **steps mean `r d[index, 2]`** 


## Imputing missing values
### Number of missing values in the dataset 
```{r echo=TRUE}
numberOfNA <- nrow(data[is.na(data$steps),])
```

- Total number of missing values in the dataset is **`r numberOfNA`**

Filling the  missing steps values by assigning the average for similar intervals.
```{r echo=TRUE}
fixed <- data
fixedNA <- is.na(fixed$steps)
fixed[fixedNA, 1] <- sapply(fixed[fixedNA, 3], function(x) dinterval[dinterval$interval == x, 2])
f <- aggregate(steps~date, fixed, sum)
plot(f$date, f$steps, type="h", ylab = 'Total steps', xlab = 'Date', 
     main = 'Total steps in 2012')
```

- The *mean* total number of steps taken per day is **`r mean(f$steps, na.rm=T)`**
- The *median* total number of steps taken per day: **`r median(f$steps, na.rm=T)`**


**What is the impact of imputing missing data on the estimates of the total daily number of steps?**  
*There is no impact on the mean total number of steps. And there is a small impact on the median.*


## Are there differences in activity patterns between weekdays and weekends?

```{r echo=TRUE}
# helper funciton to determine if a date is a weekday or a weekend
isWeekend <- function(x) {
    wd <- weekdays(x)
    if (wd %in% c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday'))
        'weekday'
    else 
        'weekend'
}

# add a new field to the dataset which indicate if the date is a weekday or a weekend
fixed$daytype <- sapply(fixed$date, isWeekend)

# calc steps mean
wdata <- aggregate(steps~interval+daytype, fixed, mean)
#weekenddata <- wdata[wdata$daytype=='weekend', c(1,3)]
#weekdaydata <- wdata[wdata$daytype=='weekday', c(1,3)]
#par(mfrow = c(2,1))
#plot(weekdaydata, type='l', ylab="Mean steps", xlab="Interval")
#plot(weekenddata, type='l', ylab="Mean steps", xlab="Interval")
library(lattice)
xyplot( steps ~ interval | daytype, wdata, type="l", xlab="Interval", ylab="Average of steps", layout=c(1,2))
```

It seems that there is more steps on weekends


